---
title: "Autoencoder on other datasets"
author: "Adrian de Luis García"
output:
  word_document: default
  html_document: default
---

This vignette has the purpose to instruct in the use of an autoencoder on different types of datasets available for study in the internet

# Environment construction

```{r, message=F, warning=F, environment}
rm(list=ls())
gc()
library(keras)
library(nnet)
library(tidyverse) # That (in)famous Mr. Wickham!
library(caret)    # To build the classifiers.
library(mlbench)  # Many databases for ML tasks
library(vcd)       # Categorical benchmarks
library(candisc)   # Wine dataset
library(entropies) # Processing and visualizing joint entropies
library(compositions)# Statistics work differently on compositional data
library(ggtern) 
library(class)
library(dplyr)

```

Let's do a simple visualization of the data involved in this analysis

```{r switches}
fancy <- TRUE  # set this for nicer on-screen visualization.
# FVA: changed to new interface: first call in the frame of datasets
data(datasets)
#datasets <- loadDataset()
datasets
splitShapesForTypes <- c("X"=4, "Y"=4, "XY"=20) #To draw split diagrams
# Naive transformation from factors to numbers in 0 to num.factors - 1
factor.as.numeric <- function(f){
  nums <- as.numeric(f)
  return(nums - min(nums))
}
```

## Choosing the Ionosphere dataset from the available options

```{r dataset choice}
dsName <- "Ionosphere"
dsRecord <-  filter(datasets, name == dsName)
# FVA: changed to use new primitive, loadDataset q.v.
ds <- loadDataset(dsRecord$name,pkg=dsRecord$packName)
#ds <- evalDataset(dsName) 

if (!is.na(dsRecord$idNumber)){
  ds <- ds[,-dsRecord$idNumber]
}
```

## Basic data from the set for classification

```{r}
#class column
ds.classNum <- which(names(ds)==dsRecord$className)
#take away the class, but keep it just in case.
class.ds <- ds[, ds.classNum]#saving the class. Warning A FACTOR!
ds <- ds[,-ds.classNum]
ds <- ds %>%     
  #transform factors to number
  mutate_if(is.factor,factor.as.numeric) %>%
  # Dispose of columns with NaN
  select_if(function(v) !any(is.na(v))) %>% 
  # Dispose of constant columns: they carry no information
  select_if(function(v)(var(v) > 0))
ncols <- ncol(ds)#Mnemonic shortcut: num of columns
dsDiscretized <- infotheo::discretize(ds, disc="equalwidth")
if (dsName != "Ionosphere"){
  log.ds <- log(ds)#this has to be made conditional on the database
  log.dsDiscretized <- infotheo::discretize(log.ds)
  #TODO: try to get rid of annoying warnings each time entropy is called. 
}
X <- as.matrix(ds)
Y <- class.ds
classes <- unique(Y)
numC <- length(classes)
print(sprintf("%s has %d classes with distribution: ", dsName, numC))
summary(Y)
```


## Design the classifier and the 5 folds with train and test 


```{r random split}


set.seed(27)
#FVA: specify #of folds
kfolds <- 5
#sample <- createDataPartition(y = Y, p=0.8, list=FALSE,times = kfolds)
sample <- createFolds(Y, k=kfolds,list=TRUE)#Tries to keep proportions in folds

#Species <- decodeClassLabels(Species)
#experiments <- 1
results <- frame()# FVA: look for appropriate syntax for empty frame

matriz <- matrix(c(0,0,0,0),nrow=2,ncol=2)
teCM_total_auto_mlp <- as.table(matriz)
teCM_total_auto_knn <- as.table(matriz)
teCM_total_pca_knn <- as.table(matriz)
 teCM_total_pca_mlp <- as.table(matriz)

accuracy_pca <- matrix(1:5,nrow = 5,ncol = 1)
accuracy_auto <- matrix(1:5,nrow = 5,ncol = 1)
for(i in (1:kfolds)){# i selects the active fold in "sample"
    # FVA: cambiar train por test
  set.seed(67)
  lm <- paste("I am on the compilation number ",i)
  lm
  trainX <- X[-sample[[i]],]
  trainY <- Y[-sample[[i]]]
  testX <- X[sample[[i]],]
  testY <- Y[sample[[i]]]
  
  hot_label_test <- toNumericClassLabels(testY)
  hot_label_test <- to_categorical(hot_label_test)
  hot_label_test <- hot_label_test[,-1]
  
  hot_label_ <- toNumericClassLabels(trainY)
  hot_label <- to_categorical(hot_label_)
  hot_label <- hot_label[,-1]
  
  batch_size <- 128
  num_classes <- 2
 
  
  encoding_dim33 <- 33
  encoding_dim50 <- 50
  encoding_dim20 <- 20
  
  input_img <- layer_input(shape = c(33))
  
  encoded_input <- layer_dense(unit = encoding_dim50,activation='relu',input_img)
  
  encoded <- layer_dense(unit = encoding_dim20,activation = 'relu',encoded_input)
  
  encoded <- layer_dense(unit = 8,activation = 'relu',encoded)
  
  
  
  decoded_input <- layer_dense(unit = encoding_dim20,activation='relu',encoded)
  
  decoded <- layer_dense(unit = encoding_dim50,activation = 'relu',decoded_input)
  
  decoded <- layer_dense(unit = encoding_dim33,activation = 'sigmoid',decoded)
  
  # FVA: I do not think this defines an AutoEncoder architecture, but 
  # only the encoder. 
  autoencoder <- keras_model(input = input_img, output = decoded)
  
  
  # FVA: falta el encoder!
  encoder <- keras_model(input = input_img, output = encoded)
  
  # FVA: therefore, this is trying to encode the input X into the intermediate layer and then to decode it with the layer "encoded"
  autoencoder %>% compile(
    optimizer='adam',
    #loss='binary_crossentropy',
    loss ='mean_squared_error',
    metrics = 'accuracy'
  )
  
  history <- autoencoder %>% fit(
    trainX,trainX,
    epochs=30,#FVA
    #batch_size=24,# FVA: select 
    steps_per_epoch=nrow(trainX),#FVA
    shuffle=TRUE
  )
  
  # FVA: this generates the X'
  
  x_train_predicted <- autoencoder %>% predict(trainX)
  x_test_predicted <- autoencoder %>% predict(testX)
  
  # FVA: this should instead generate the Z's
  # AT PRESENT THEY CANNOT BE GENERATED, because the SAE architecture is wrong.
  
  # FVA: I don't understand this rewriting: 
  # Is this trying to obtain the Zs?
  # Why systematically rewrite the 1:4 folds?
  # This should take into consideration the "i" variable for the active fold.

  # FVA: A) Predict Z
  # FVA: B) class9fy
  trainZ <- encoder %>% predict(trainX)
  testZ<- encoder %>% predict(testX)
  
  train_matrixZ <- as.matrix(trainZ)
  test_matrixZ <- as.matrix(testZ)
  
  #####################################################################################
  
  colnames(train_matrixZ) <- c("1st","2nd","3rd","4rd","5th","6th","7th","8th")
  colnames(test_matrixZ) <- c("1st","2nd","3rd","4rd","5th","6th","7th","8th")
  fit <- caret::train(x=train_matrixZ, y=trainY, 
                 method="knn",
                 preProcess = c("center","scale"),
               tuneLength = 15
               )

  trCM <- confusionMatrix(predict(fit,train_matrixZ), trainY)
  
  trCoords <- jentropies(t(trCM))
  
  teCM <- confusionMatrix(predict(fit,test_matrixZ), testY)
  
  teCoords <- jentropies(t(teCM))
  
  
  
  teCM_total_auto_knn <- teCM_total_auto_knn + teCM
  
  results <- rbind(results,
                    cbind(
                        dSet="Ionosphere", fold=i, method="knn",
                        rbind(
                            cbind(trCoords,Phase="train"), 
                            cbind(teCoords, Phase="test")
                            )
                        )
              )
  
  ###################################################################################

     pca <- prcomp(trainX) 
  
     train_pca <- predict(pca,newdata = trainX)
     test_pca <- predict(pca,newdata = testX)

     fit_pca <- caret::train(x=predict(pca,newdata = trainX), y=trainY, 
                 method="knn",
                 preProcess = c("center","scale"),
               tuneLength = 15
               )
  
  
  trCM <- confusionMatrix(predict(fit_pca,train_pca), trainY)
  
  trCoords <- jentropies(t(trCM))
  
  teCM <- confusionMatrix(predict(fit_pca,test_pca), testY)
  
  teCoords <- jentropies(t(teCM))
  
  
  
  teCM_total_pca_knn <- teCM_total_pca_knn + teCM
  
  results <- rbind(results,
                    cbind(
                        dSet="Ionosphere", fold=i, method="knn_pca",
                        rbind(
                            cbind(trCoords,Phase="train"), 
                            cbind(teCoords, Phase="test")
                            )
                        )
              )
  
  
# train_matrixZ <- normalize(x = train_matrixZ, method = "range", range = c(0, 1))
# FVA: summaries of cross plots
mlp_pca <- keras_model_sequential()
mlp_pca %>% 
  layer_dense(units = 40, activation = 'relu', input_shape = c(33)) %>% 
  layer_dense(units = 20, activation = 'relu') %>%
  layer_dense(units = 10, activation = 'sigmoid') %>%
  layer_dense(units = 2, activation = 'softmax')

    mlp_pca %>% compile(
    loss = 'binary_crossentropy',
    optimizer = 'adam',
    metrics = c('accuracy')
  )
  
  history <- mlp_pca %>% fit(
    train_pca, hot_label,
    batch_size = batch_size,
    epochs = 40
  )

  accuracy_pca[i] <- max(history$metrics$acc)
                           
  trCM <- confusionMatrix(predict(mlp_pca,train_pca), hot_label)
  
  trCoords <- jentropies(t(trCM))
  
  teCM <- confusionMatrix(predict(mlp_pca,test_pca),hot_label_test)
  
  teCoords <- jentropies(t(teCM))
  
 
  
   teCM_total_pca_mlp <- teCM_total_pca_mlp + teCM
   
  results <- rbind(results,
                    cbind(
                        dSet="Ionosphere", fold=i, method="mlp_pca",
                        rbind(
                            cbind(trCoords,Phase="train"), 
                            cbind(teCoords, Phase="test")
                            )
                        )
                )
  
  
  
  
############################################################################################################################################

  
  hot_label_test <- toNumericClassLabels(testY)
  hot_label_ <- toNumericClassLabels(trainY)
  hot_label <- to_categorical(hot_label_)
  hot_label <- hot_label[,-1]
  
  summary(train_matrixZ)
# train_matrixZ <- normalize(x = train_matrixZ, method = "range", range = c(0, 1))
# FVA: summaries of cross plots
mlp <- keras_model_sequential()
mlp %>% 
  layer_dense(units = 12, activation = 'relu', input_shape = c(8)) %>% 
  layer_dense(units = 6, activation = 'sigmoid') %>%
  layer_dense(units = 2, activation = 'softmax')

#  input_mlp <- layer_input(shape = c(4))
  
#  input_layer_mlp <- layer_dense(unit = 5,activation='relu',input_mlp)
  
#  middle_layer_mlp <- layer_dense(unit = 4,activation = 'sigmoid',input_layer_mlp)
  
#  output_layer_mlp <- layer_dense(unit = 3,activation = 'sigmoid',middle_layer_mlp) 
   
#  mlp <- keras_model(input = input_mlp, output = output_layer_mlp)
  
  mlp %>% compile(
    loss = 'binary_crossentropy',
    optimizer = 'adam',
    metrics = c('accuracy')
  )
  
  history <- mlp %>% fit(
    train_matrixZ, hot_label,
    batch_size = batch_size,
    epochs = 100
  )
  
  accuracy_auto[i] <- max(history$metrics$acc)
  
  test_mlp <- predict(mlp,train_matrixZ)
  
  #trCM <- confusionMatrix(predict(mlp,train_matrixZ), hot_label_)
  
  #trCoords <- jentropies(t(trCM))
      
                         
  trCM <- confusionMatrix(predict(mlp,train_matrixZ), hot_label)
  
  trCoords <- jentropies(t(trCM))
  
  teCM <- confusionMatrix(predict(mlp,test_matrixZ),hot_label_test)
  
  teCoords <- jentropies(t(teCM))
  
  
  
  teCM_total_auto_mlp <- teCM_total_auto_mlp + teCM
  
  results <- rbind(results,
                    cbind(
                        dSet="Ionosphere", fold=i, method="mlp",
                        rbind(
                            cbind(trCoords,Phase="train"), 
                            cbind(teCoords, Phase="test")
                            )
                        )
                )
 
}
```

## Including Plots from the first autoencoder and the mlp

You can also embed plots, for example:

```{r plots}
Total_Resultado_mlp <- frame()
Total_Resultado_mlp <- rbind(Total_Resultado_mlp,
                    cbind(
                         dSet="Ionosphere", method="mlp",
                        rbind(
                            cbind(jentropies(t(teCM_total_pca_mlp)), Classifier="pca")
                            )
                        )
                )
Total_Resultado_mlp <- rbind(Total_Resultado_mlp,
                    cbind(
                         dSet="Ionosphere", method="mlp",
                        rbind(
                            cbind(jentropies(t(teCM_total_auto_mlp)),Classifier="auto")
                            )
                        )
                )

Total_Resultado_mlp <- rbind(Total_Resultado_mlp,
                    cbind(
                         dSet="Ionosphere", method="knn",
                        rbind(
                            cbind(jentropies(t(teCM_total_pca_knn)), Classifier="pca")
                            )
                        )
                )

Total_Resultado_mlp <- rbind(Total_Resultado_mlp,
                    cbind(
                         dSet="Ionosphere", method="knn",
                        rbind(
                           cbind(jentropies(t(teCM_total_auto_knn)),Classifier="auto")
                            )
                        )
                )
gp_Total <- ggmetern(ed = (Total_Resultado_mlp %>% filter(type == "XY")), fancy = TRUE) +
    geom_point(aes(colour=Classifier, shape = method), size=2) + labs(shape="Methods") + 
    scale_colour_brewer(palette="Set1")


gp <- ggmetern(ed= (results %>% filter(type == "XY") %>% filter(method == "knn") %>% filter(dSet == "Ionosphere")) , fancy = TRUE) +
  geom_point(aes(colour=Phase, shape=as.character(fold)), size=2)  +
  labs(shape="Fold") + 
  scale_colour_brewer(palette="Set1")
gp1 <- ggmetern(ed= (results %>% filter(type == "XY") %>% filter(method == "mlp") %>% filter(dSet == "Ionosphere")) , fancy = TRUE) +
  geom_point(aes(colour=Phase, shape=as.character(fold)), size=2)  +
  labs(shape="Fold") + 
  scale_colour_brewer(palette="Set1")
gp3 <- ggmetern(ed= (results %>% filter(type == "XY") %>% filter(method == "knn_pca") %>% filter(dSet == "Ionosphere")) , fancy = TRUE) +
  geom_point(aes(colour=Phase, shape=as.character(fold)), size=2)  +
  labs(shape="Fold") + 
  scale_colour_brewer(palette="Set1")

gp4 <- ggmetern(ed= (results %>% filter(type == "XY") %>% filter(method == "mlp_pca") %>% filter(dSet == "Ionosphere")) , fancy = TRUE) +
  geom_point(aes(colour=Phase, shape=as.character(fold)), size=2)  +
  labs(shape="Fold") + 
  scale_colour_brewer(palette="Set1")
gp

gp1

gp3

gp4

gp_Total

```

